---
title: "Data Wrangling Exercise 1"
output: html_document
---
##0: Load the data in RStudio
Save the data set as a CSV file called refine_original.csv and load it in RStudio into a data frame.
```{datacsv}
#load libraries
library(dplyr)
library(stringr)
library(tidyr)

#load csv as dataframe
datacsv <- read.csv("/users/Jessie/Downloads/refine_original.csv")

```

## 1: Clean up brand names
Clean up the 'company' column, so all of the misspellings of the brand names are standardized. For example, you can transform the values in the column to be: philips, akzo, van houten and unilever (all lowercase).

```{r}
#Look for the unique variables in company
unique(datacsv$company)

#Make all names lowercase
datacsv$company <- tolower(datacsv$company)

#Replace misspelled names with philips, akzo, van houten, or unilever
datacsv$company <- str_replace(datacsv$company, "akz0|ak zo", "akzo")
datacsv$company <- str_replace(datacsv$company, "phillips|phllips|phillps|fillips|phlips", "philips")
datacsv$company <- str_replace(datacsv$company, "unilver", "unilever")

```

## 2: Separate product code and number
Separate the product code and product number into separate columns i.e. add two new columns called product_code and product_number, containing the product code and number respectively

```{r}
#Product.code...number will be separated into Product_Code and Product_Number
datacsv <- separate(datacsv, Product.code...number, c("product_code", "product_number"), sep = "-")

#Remove Product.code...number
datacsv$Product.code...number <- NULL

```

## 3: Add product categories
You learn that the product codes actually represent the following product categories:
p = Smartphone
v = TV
x = Laptop
q = Tablet
In order to make the data more readable, add a column with the product category for each record.

```{refine$Product_Category}
#Add product_category
product_category <- datacsv$product_code
datacsv <- cbind(datacsv, product_category)

#define product codes
datacsv$product_category <- str_replace(datacsv$product_category, "p", "Smartphone")
datacsv$product_category <- str_replace(datacsv$product_category, "v", "TV")
datacsv$product_category <- str_replace(datacsv$product_category, "x", "Laptop")
datacsv$product_category <- str_replace(datacsv$product_category, "q", "Tablet")

```

## 4: Add full address for geocoding
You'd like to view the customer information on a map. In order to do that, the addresses need to be in a form that can be easily geocoded. Create a new column full_address that concatenates the three address fields (address, city, country), separated by commas.

```{r}
#Unite the address, city, and country columns into one
datacsv <- unite(datacsv, Full_Address, address, city, country, sep =",")

#Remove address, city, country columns
datacsv$address <- NULL
datacsv$city <- NULL
datacsv$country <- NULL

```

## 5: Create dummy variables for company and product category
Both the company name and product category are categorical variables i.e. they take only a fixed set of values. In order to use them in further analysis you need to create dummy variables. Create dummy binary variables for each of them with the prefix company_ and product_ i.e.
Add four binary (1 or 0) columns for company: company_philips, company_akzo, company_van_houten and company_unilever
Add four binary (1 or 0) columns for product category: product_smartphone, product_tv, product_laptop and product_tablet

```{r}
#create binary columns for company
company_akzo <- as.numeric(datacsv$company == 'akzo')
company_philips <- as.numeric(datacsv$company == 'philips')
company_van_houten <- as.numeric(datacsv$company == "van houten")
company_unilever <- as.numeric(datacsv$company == 'unilever') 

#bind new company binary columns to dataframe
datacsv <- cbind(datacsv, company_akzo, company_philips, company_van_houten, company_unilever)

#create binary columns for product_category
product_smartphone <- as.numeric(datacsv$product_category == 'Smartphone')
product_tv <- as.numeric(datacsv$product_category == 'TV')
product_laptop <- as.numeric(datacsv$product_category == 'Laptop')
product_tablet <- as.numeric(datacsv$product_category == 'Tablet')

#bind new product_category binary columns to dataframe
datacsv <- cbind(datacsv, product_smartphone, product_tv, product_laptop, product_tablet)

```

##6: Submit the project on Github
Include your code, the original data as a CSV file refine_original.csv, and the cleaned up data as a CSV file called refine_clean.csv.

```{r}
#Convert file to csv file
write.csv(datacsv, file = "refine_clean.csv")

```